Pulmonary acini exhibit complex changes during postnatal rat lung development

### 2021-09-24 manuscript.tex
PLOS ONE wants *all* data available [ci skip]
This build is based on
https://github.com/habi/acinar-analysis-manuscript/commit/f456264a8870c9b4324894a3ec64a04fcf8280f3.

This commit was created by the following CI build and job:
https://github.com/habi/acinar-analysis-manuscript/commit/f456264a8870c9b4324894a3ec64a04fcf8280f3/checks
https://github.com/habi/acinar-analysis-manuscript/runs/1269084033

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681}.
It has been accepted for publication in \href{https://journals.plos.org/plosone/}{\emph{PLOS ONE}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/5e778036e316c47b8bac4ce50086b191a7e3707a...0e2445c24e13ea304a7439efa969e7a0fc2309cd}{has been updated} since submission.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/0e2445c24e13ea304a7439efa969e7a0fc2309cd/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/0e2445c24e13ea304a7439efa969e7a0fc2309cd}{habi/acinar-analysis-manuscript@0e2445c}
on September 10, 2021.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(0e2445c)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(0e2445c)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\end{itemize}

@@

Previously, we have shown that the stereological assessment of the acinar volume (according to the Cavalieri principle {[}\protect\hyperlink{ref-FE9HLB4f}{36}{]}) gives comparable results to an automated volume extraction by counting the segmented pixels in the tomographic data sets {[}\protect\hyperlink{ref-7YLeeyu}{21}{]}.
Due to variations in the gray value, the automatic segmentation underestimates the volume of the single acini.
The best-suited approach to obtain unbiased results is to assess the volume manually according to the Cavalieri principle, which is what we did for the 268 acini presented in this manuscript.
The sets of JPG slices and the raw results from the stereological analysis with the STEPanizer are available on request.

@@

All the stereologically assessed data was processed in a \href{http://jupyter.org/}{Jupyter} {[}\protect\hyperlink{ref-pQ6Wbz73}{37}{]} notebook, producing all the results, including the statistical data and plots shown below.
The notebook with all its history is freely available on GitHub {[}\protect\hyperlink{ref-V87xbt0b}{38}{]}.
The performed calculations are described in detail in the \protect\hyperlink{results}{Results} section.
In the plots, semitransparent circles mark the single observations.
The box shows the 25~\%--75~\% quartiles range of the data.
The whiskers extend to the rest of the distribution.
Outliers are determined as a function of the inter-quartile range and are shown outside the whiskers.
Numerical values in the text are given as averages ± standard deviation.
P-values in the text and figure legends are given as precise numbers, as suggested by Amrhein et al.~{[}\protect\hyperlink{ref-o21zxPIu}{39}{]}.

Usually, we performed a Shapiro-Wilk test for normality {[}\protect\hyperlink{ref-IkHrgIj3}{40}{]} to test whether we can use an U-test for assessing the differences, namely a two-sided Mann-Whitney rank test {[}\protect\hyperlink{ref-MyzxBJ57}{41}{]}.
This rank test was used to assess the differences between the possible combinations.
An additional Kruskal-Wallis H-test for independent samples {[}\protect\hyperlink{ref-ON1Bppkk}{42}{]} was used to test for sample independence.
The statistical analysis was performed in the aforementioned notebook by using the statistical functions of the Scientific Computing Tools for Python {[}\protect\hyperlink{ref-8Miti2Gz}{43}{]} or GraphPad Prism 7.01 (GraphPad Software, San Diego, CA, USA).

@@

The alveolar number shows highly significant differences (Šidák-corrected p-value smaller than 0.00167 {[}\protect\hyperlink{ref-GtCynqsI}{44}{]}) between all possible combination of days.
All possible combinations of entrance ring counts per day are significantly different (all p-values are smaller than p=1.9e-5, which is the p-value of the difference between days 4 and 10).
The entrance ring counts for all animals are independent (p=1.6e-34).

@@

We can easily estimate the total number of acini during lung development from our data.
The limitations of this study are mainly twofold: a) due to technical reasons (i.e.~available beamtime and limitations of the setup) we were not able to image the entire lung and b) lung location (i.e.~lower medial tip of the right lobe).
However, our sample is more than suitable for the drawn conclusions because it was previously shown that the right lower lobe is a valid sample for the entire lung {[}\protect\hyperlink{ref-14OP85b2F}{10}{]} and that information drawn from one lobe is representative of the whole lung {[}\protect\hyperlink{ref-I9TmP6IU}{45}{]}.
Furthermore, based on stereological data, Zeltner et al.~{[}\protect\hyperlink{ref-I9TmP6IU}{45}{]} were able to show that the lung parenchyma is rather homogeneous.
Looking at individual acini, Kizhakke Puliyakote et al.~{[}\protect\hyperlink{ref-RGBeCf8v}{22}{]} observed that the peripheral, pleura facing acini possess an acinar volume which is two thirds larger than the volume of the central ones.
However, they did not report additional regional differences, for example between acini located cranially and caudally.
Own preliminary data point to the same result.
We imaged an entire rat lung \emph{in situ} immediately post mortem using high-resolution synchrotron radiation based X-ray tomographic microscopy {[}\protect\hyperlink{ref-Zv1XcgAO}{46}{]}.
A first analysis of regional difference shows similar results as mentioned above.

@@

By dividing the parenchymal volumes of each lung estimated by Tschanz et al.~(Table 1, {[}\protect\hyperlink{ref-wnl86DEM}{6}{]}) by the mean number of acini estimated by Barré et al.~{[}\protect\hyperlink{ref-uFNlWogb}{11}{]} we can estimate the mean acinar volume for each day.
The acinus volume we estimated by point counting (Cavalieri estimation) are on average 2.07 times lower but show a similar increase over the studied period (see figure \ref{fig:06}).
It appears that the comparison between data obtained by stereology based on classical paraffin sections or on any kind of 3D-imaging has its limitation if it comes to the absolute numbers.
First, due to the \emph{coast of Britain}-effect the exact value of every surface measurement is depending on the resolution of the imaging method {[}\protect\hyperlink{ref-AdXhaEV4}{47}{]}.
Second, the thickness of the alveolar septa in 3D-imaging data is heavily depending on the segmentation between airspace and tissue---regardless if it is done manually or by a computed algorithm.
We observed that the airspace volumes obtained by 3D-imaging has a tendency to be smaller than the ones determined using classical paraffin sections.
Nonetheless, the relative numbers between studies to match well.

@@

The volume and shape of the alveoli, alveolar ducts, and the acini are critical parameters for ventilation and particle deposition {[}\protect\hyperlink{ref-eioib1TQ}{48}--\protect\hyperlink{ref-OT4s1CSX}{50}{]}.
Particle deposition, for example, is low during the first two weeks of a rats' life, high at postnatal day P21 and medium at days P36--P90 {[}\protect\hyperlink{ref-yHHhvOtP}{15},\protect\hyperlink{ref-18DcNLAv6}{16},\protect\hyperlink{ref-VJYqPtdT}{51}{]}.
This correlates well with our results, where we found a significantly larger number of alveoli per acinus volume at days 10 and 21 compared to days 4 and 60 (all p-values are smaller than 5.4e-5, which is the p-value found between day 4 and~day 21).
Kreyling et al.~{[}\protect\hyperlink{ref-18DcNLAv6}{16}{]} report maximal gold nanoparticle retention at this time point (see \href{https://pubs.acs.org/doi/10.1021/acsnano.8b01826\#fig1}{their figure 1}).

@@

As shown in a study by Osmanagic et al.~{[}\protect\hyperlink{ref-LzalnLBH}{52}{]} the total alveolar number shows a large variability between different labs.
E.g.~the reported total number of alveoli in adult C57BL6 mice differs by a factor of approximately 8.
All the studies mentioned by Osmanagic et al.~were performed in labs well-known for their quality of stereology-based investigations of lung morphology.
Therefore, any simple bias or methodical error may not account for the reported differences.
It seems to be that the total number of alveoli is dependent on unknown small methodical differences and minute differences in how the operator actually does count.
The seemingly minuscule volume fraction of lung studied (around 1 \% or less) in this manuscript is---according to the stereological guidelines {[}\protect\hyperlink{ref-dNc8FfNn}{30}{]}---ample volume to gain mathematically correct results if the stereology is carried out correctly as we have done to the best of our knowledge.
For human lungs for example a so-called sampling cascade is used to study details down to the thickness of the air-blood barrier (0.5 µm, see Figure 1 of {[}\protect\hyperlink{ref-1CLJ32t5m}{53}{]} for a visualization of this cascade).
The studied volume is in that case multiple decimal powers smaller than what is studied here while still being mathematically sound and accepted.

However, relative numbers as reported in this study are more reliable.
This is especially the case when the comparison is done in groups of animals processed in parallel and counted by the same operator.
We assessed the same lungs as Tschanz et al.~{[}\protect\hyperlink{ref-wnl86DEM}{6}{]}, but performed our assessment based on a different imaging method and with two different persons performing the stereological assessment at two different points in time.
Furthermore, we used the number of acini as determined by Barré et al.~{[}\protect\hyperlink{ref-uFNlWogb}{11}{]} which represents a dataset obtained by a third team and which introduces the effect of error propagation.
This may explain the relative factor of 5.04 that we found as a difference.
Even if it was shown that the right lower lobe represents a valid sample for the entire lung {[}\protect\hyperlink{ref-14OP85b2F}{10},\protect\hyperlink{ref-I9TmP6IU}{45}{]}, we cannot rule out a contribution to the factor by regional differences in the lung which we cannot account for in the present study.

@@

It has been predicted by computational fluid dynamics simulations and by to-scale experiments that the amount and location of massless particle deposition, as well as ventilation, are depending on the size of the acini {[}\protect\hyperlink{ref-CVl41LwO}{12},\protect\hyperlink{ref-1DP2FRUSZ}{14},\protect\hyperlink{ref-RJ9GQcwl}{54}{]}.
The predictions could be confirmed at least in one case, where mice received fluorescent particles during mechanical ventilation.
In 3D-visualizations the particles were predominately detected in the proximal half of the acini {[}\protect\hyperlink{ref-19jBv0ima}{55}{]}.

Based on the above-mentioned knowledge and the data of this study, we propose the following.
In the smallest acini we observed at postnatal day 4 acinar flows are mainly characterized by radial streamlines {[}\protect\hyperlink{ref-eioib1TQ}{48},\protect\hyperlink{ref-afF83siN}{56}{]}.
Diffusion governs massless particle deposition {[}\protect\hyperlink{ref-k5fqLjiy}{57}{]}.
As the acini increase in size, convective transport becomes more and more dominant in the proximal regions of the acini.
In larger acini, the proximal region shows much higher particle retention than the distal one, because in this region the larger air-flow induces a circular current, which facilitates alveolar particle deposition.
The proximal regions may even be viewed analogously to a filter capturing the particles whereas the distal regions do not receive significant amounts {[}\protect\hyperlink{ref-CVl41LwO}{12},\protect\hyperlink{ref-1DP2FRUSZ}{14},\protect\hyperlink{ref-yHHhvOtP}{15},\protect\hyperlink{ref-1HOyVjqpM}{49},\protect\hyperlink{ref-OT4s1CSX}{50},\protect\hyperlink{ref-RJ9GQcwl}{54}{]}.

@@

We thank \texttt{manubot} {[}\protect\hyperlink{ref-YuJbg3zO}{58}{]} for providing us with the means to collaboratively write and craft this manuscript.

@@

\leavevmode\hypertarget{ref-pQ6Wbz73}{}%
37. Kluyver T, Ragan-Kelley B, Pérez F, Granger B, Bussonnier M, Frederic J, et al. Jupyter Notebooks -- a publishing format for reproducible computational workflows. In: Loizides F, Scmidt B, editors. IOS Press; 2016. pp. 87--90. doi:\href{https://doi.org/10.3233/978-1-61499-649-1-87}{10.3233/978-1-61499-649-1-87}

\leavevmode\hypertarget{ref-V87xbt0b}{}%
38. Haberthür D. habi/acinar-analysis: AJP-Submission. Zenodo; 2020. doi:\href{https://doi.org/10.5281/zenodo.3768407}{10.5281/zenodo.3768407}

\leavevmode\hypertarget{ref-o21zxPIu}{}%
39. Amrhein V, Greenland S, McShane B. Scientists rise up against statistical significance. Nature. 2019;567: 305--307. doi:\href{https://doi.org/10.1038/d41586-019-00857-9}{10.1038/d41586-019-00857-9}

\leavevmode\hypertarget{ref-IkHrgIj3}{}%
40. SHAPIRO SS, WILK MB. An analysis of variance test for normality (complete samples). Biometrika. 1965;52: 591--611. doi:\href{https://doi.org/10.1093/biomet/52.3-4.591}{10.1093/biomet/52.3-4.591}

\leavevmode\hypertarget{ref-MyzxBJ57}{}%
41. Mann HB, Whitney DR. On a Test of Whether one of Two Random Variables is Stochastically Larger than the Other. The Annals of Mathematical Statistics. 1947;18: 50--60. doi:\href{https://doi.org/10.1214/aoms/1177730491}{10.1214/aoms/1177730491}

\leavevmode\hypertarget{ref-ON1Bppkk}{}%
42. Kruskal WH, Wallis WA. Use of Ranks in One-Criterion Variance Analysis. Journal of the American Statistical Association. 1952;47: 583--621. doi:\href{https://doi.org/10.1080/01621459.1952.10483441}{10.1080/01621459.1952.10483441}

\leavevmode\hypertarget{ref-8Miti2Gz}{}%
43. Virtanen P, Gommers R, Oliphant TE, Haberland M, Reddy T, Cournapeau D, et al. SciPy 1.0: fundamental algorithms for scientific computing in Python. Nature Methods. 2020;17: 261--272. doi:\href{https://doi.org/10.1038/s41592-019-0686-2}{10.1038/s41592-019-0686-2}

\leavevmode\hypertarget{ref-GtCynqsI}{}%
44. Šidák Z. Rectangular Confidence Regions for the Means of Multivariate Normal Distributions. Journal of the American Statistical Association. 1967;62: 626--633. doi:\href{https://doi.org/10.1080/01621459.1967.10482935}{10.1080/01621459.1967.10482935}

\leavevmode\hypertarget{ref-I9TmP6IU}{}%
45. Zeltner TB, Bertacchini M, Messerli A, Burri PH. Morphometric Estimation of Regional Differences in the Rat Lung. Experimental Lung Research. 2009;16: 145--158. doi:\href{https://doi.org/10.3109/01902149009087879}{10.3109/01902149009087879}

\leavevmode\hypertarget{ref-Zv1XcgAO}{}%
46. Borisova E, Lovric G, Miettinen A, Fardin L, Bayat S, Larsson A, et al. Micrometer-resolution X-ray tomographic full-volume reconstruction of an intact post-mortem juvenile rat lung. Histochemistry and Cell Biology. 2020;155: 215--226. doi:\href{https://doi.org/10.1007/s00418-020-01868-8}{10.1007/s00418-020-01868-8}

\leavevmode\hypertarget{ref-AdXhaEV4}{}%
47. Mandelbrot B. How Long Is the Coast of Britain? Statistical Self-Similarity and Fractional Dimension. Science. 1967;156: 636--638. doi:\href{https://doi.org/10.1126/science.156.3775.636}{10.1126/science.156.3775.636}

\leavevmode\hypertarget{ref-eioib1TQ}{}%
48. Sznitman J, Sutter R, Altorfer D, Stampanoni M, Rösgen T, Schittny JC. Visualization of respiratory flows from 3D reconstructed alveolar airspaces using X-ray tomographic microscopy. Journal of Visualization. 2010;13: 337--345. doi:\href{https://doi.org/10.1007/s12650-010-0043-0}{10.1007/s12650-010-0043-0}

\leavevmode\hypertarget{ref-1HOyVjqpM}{}%
49. Henry FS, Haber S, Haberthür D, Filipovic N, Milasinovic D, Schittny JC, et al. The Simultaneous Role of an Alveolus as Flow Mixer and Flow Feeder for the Deposition of Inhaled Submicron Particles. Journal of Biomechanical Engineering. 2012;134: 121001. doi:\href{https://doi.org/10.1115/1.4007949}{10.1115/1.4007949}

>>>>

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681}.
It has been accepted for publication in \href{https://journals.plos.org/plosone/}{\emph{PLOS ONE}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/5e778036e316c47b8bac4ce50086b191a7e3707a...f456264a8870c9b4324894a3ec64a04fcf8280f3}{has been updated} since submission.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/f456264a8870c9b4324894a3ec64a04fcf8280f3/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/f456264a8870c9b4324894a3ec64a04fcf8280f3}{habi/acinar-analysis-manuscript@f456264}
on September 24, 2021.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(f456264)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(f456264)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\end{itemize}

@@

Previously, we have shown that the stereological assessment of the acinar volume (according to the Cavalieri principle {[}\protect\hyperlink{ref-FE9HLB4f}{36}{]}) gives comparable results to an automated volume extraction by counting the segmented pixels in the tomographic data sets {[}\protect\hyperlink{ref-7YLeeyu}{21}{]}.
Due to variations in the gray value, the automatic segmentation underestimates the volume of the single acini.
The best-suited approach to obtain unbiased results is to assess the volume manually according to the Cavalieri principle, which is what we did for the 268 acini presented in this manuscript.
The sets of JPG slices and the raw results from the stereological analysis with the STEPanizer are available online {[}\protect\hyperlink{ref-14ETRiHY2}{37}{]}.

@@

All the stereologically assessed data was processed in a \href{http://jupyter.org/}{Jupyter} {[}\protect\hyperlink{ref-pQ6Wbz73}{38}{]} notebook, producing all the results, including the statistical data and plots shown below.
The notebook with all its history is freely available on GitHub {[}\protect\hyperlink{ref-V87xbt0b}{39}{]}.
The performed calculations are described in detail in the \protect\hyperlink{results}{Results} section.
In the plots, semitransparent circles mark the single observations.
The box shows the 25~\%--75~\% quartiles range of the data.
The whiskers extend to the rest of the distribution.
Outliers are determined as a function of the inter-quartile range and are shown outside the whiskers.
Numerical values in the text are given as averages ± standard deviation.
P-values in the text and figure legends are given as precise numbers, as suggested by Amrhein et al.~{[}\protect\hyperlink{ref-o21zxPIu}{40}{]}.

Usually, we performed a Shapiro-Wilk test for normality {[}\protect\hyperlink{ref-IkHrgIj3}{41}{]} to test whether we can use an U-test for assessing the differences, namely a two-sided Mann-Whitney rank test {[}\protect\hyperlink{ref-MyzxBJ57}{42}{]}.
This rank test was used to assess the differences between the possible combinations.
An additional Kruskal-Wallis H-test for independent samples {[}\protect\hyperlink{ref-ON1Bppkk}{43}{]} was used to test for sample independence.
The statistical analysis was performed in the aforementioned notebook by using the statistical functions of the Scientific Computing Tools for Python {[}\protect\hyperlink{ref-8Miti2Gz}{44}{]} or GraphPad Prism 7.01 (GraphPad Software, San Diego, CA, USA).

@@

The alveolar number shows highly significant differences (Šidák-corrected p-value smaller than 0.00167 {[}\protect\hyperlink{ref-GtCynqsI}{45}{]}) between all possible combination of days.
All possible combinations of entrance ring counts per day are significantly different (all p-values are smaller than p=1.9e-5, which is the p-value of the difference between days 4 and 10).
The entrance ring counts for all animals are independent (p=1.6e-34).

@@

We can easily estimate the total number of acini during lung development from our data.
The limitations of this study are mainly twofold: a) due to technical reasons (i.e.~available beamtime and limitations of the setup) we were not able to image the entire lung and b) lung location (i.e.~lower medial tip of the right lobe).
However, our sample is more than suitable for the drawn conclusions because it was previously shown that the right lower lobe is a valid sample for the entire lung {[}\protect\hyperlink{ref-14OP85b2F}{10}{]} and that information drawn from one lobe is representative of the whole lung {[}\protect\hyperlink{ref-I9TmP6IU}{46}{]}.
Furthermore, based on stereological data, Zeltner et al.~{[}\protect\hyperlink{ref-I9TmP6IU}{46}{]} were able to show that the lung parenchyma is rather homogeneous.
Looking at individual acini, Kizhakke Puliyakote et al.~{[}\protect\hyperlink{ref-RGBeCf8v}{22}{]} observed that the peripheral, pleura facing acini possess an acinar volume which is two thirds larger than the volume of the central ones.
However, they did not report additional regional differences, for example between acini located cranially and caudally.
Own preliminary data point to the same result.
We imaged an entire rat lung \emph{in situ} immediately post mortem using high-resolution synchrotron radiation based X-ray tomographic microscopy {[}\protect\hyperlink{ref-Zv1XcgAO}{47}{]}.
A first analysis of regional difference shows similar results as mentioned above.

@@

By dividing the parenchymal volumes of each lung estimated by Tschanz et al.~(Table 1, {[}\protect\hyperlink{ref-wnl86DEM}{6}{]}) by the mean number of acini estimated by Barré et al.~{[}\protect\hyperlink{ref-uFNlWogb}{11}{]} we can estimate the mean acinar volume for each day.
The acinus volume we estimated by point counting (Cavalieri estimation) are on average 2.07 times lower but show a similar increase over the studied period (see figure \ref{fig:06}).
It appears that the comparison between data obtained by stereology based on classical paraffin sections or on any kind of 3D-imaging has its limitation if it comes to the absolute numbers.
First, due to the \emph{coast of Britain}-effect the exact value of every surface measurement is depending on the resolution of the imaging method {[}\protect\hyperlink{ref-AdXhaEV4}{48}{]}.
Second, the thickness of the alveolar septa in 3D-imaging data is heavily depending on the segmentation between airspace and tissue---regardless if it is done manually or by a computed algorithm.
We observed that the airspace volumes obtained by 3D-imaging has a tendency to be smaller than the ones determined using classical paraffin sections.
Nonetheless, the relative numbers between studies to match well.

@@

The volume and shape of the alveoli, alveolar ducts, and the acini are critical parameters for ventilation and particle deposition {[}\protect\hyperlink{ref-eioib1TQ}{49}--\protect\hyperlink{ref-OT4s1CSX}{51}{]}.
Particle deposition, for example, is low during the first two weeks of a rats' life, high at postnatal day P21 and medium at days P36--P90 {[}\protect\hyperlink{ref-yHHhvOtP}{15},\protect\hyperlink{ref-18DcNLAv6}{16},\protect\hyperlink{ref-VJYqPtdT}{52}{]}.
This correlates well with our results, where we found a significantly larger number of alveoli per acinus volume at days 10 and 21 compared to days 4 and 60 (all p-values are smaller than 5.4e-5, which is the p-value found between day 4 and~day 21).
Kreyling et al.~{[}\protect\hyperlink{ref-18DcNLAv6}{16}{]} report maximal gold nanoparticle retention at this time point (see \href{https://pubs.acs.org/doi/10.1021/acsnano.8b01826\#fig1}{their figure 1}).

@@

As shown in a study by Osmanagic et al.~{[}\protect\hyperlink{ref-LzalnLBH}{53}{]} the total alveolar number shows a large variability between different labs.
E.g.~the reported total number of alveoli in adult C57BL6 mice differs by a factor of approximately 8.
All the studies mentioned by Osmanagic et al.~were performed in labs well-known for their quality of stereology-based investigations of lung morphology.
Therefore, any simple bias or methodical error may not account for the reported differences.
It seems to be that the total number of alveoli is dependent on unknown small methodical differences and minute differences in how the operator actually does count.
The seemingly minuscule volume fraction of lung studied (around 1 \% or less) in this manuscript is---according to the stereological guidelines {[}\protect\hyperlink{ref-dNc8FfNn}{30}{]}---ample volume to gain mathematically correct results if the stereology is carried out correctly as we have done to the best of our knowledge.
For human lungs for example a so-called sampling cascade is used to study details down to the thickness of the air-blood barrier (0.5 µm, see Figure 1 of {[}\protect\hyperlink{ref-1CLJ32t5m}{54}{]} for a visualization of this cascade).
The studied volume is in that case multiple decimal powers smaller than what is studied here while still being mathematically sound and accepted.

However, relative numbers as reported in this study are more reliable.
This is especially the case when the comparison is done in groups of animals processed in parallel and counted by the same operator.
We assessed the same lungs as Tschanz et al.~{[}\protect\hyperlink{ref-wnl86DEM}{6}{]}, but performed our assessment based on a different imaging method and with two different persons performing the stereological assessment at two different points in time.
Furthermore, we used the number of acini as determined by Barré et al.~{[}\protect\hyperlink{ref-uFNlWogb}{11}{]} which represents a dataset obtained by a third team and which introduces the effect of error propagation.
This may explain the relative factor of 5.04 that we found as a difference.
Even if it was shown that the right lower lobe represents a valid sample for the entire lung {[}\protect\hyperlink{ref-14OP85b2F}{10},\protect\hyperlink{ref-I9TmP6IU}{46}{]}, we cannot rule out a contribution to the factor by regional differences in the lung which we cannot account for in the present study.

@@

It has been predicted by computational fluid dynamics simulations and by to-scale experiments that the amount and location of massless particle deposition, as well as ventilation, are depending on the size of the acini {[}\protect\hyperlink{ref-CVl41LwO}{12},\protect\hyperlink{ref-1DP2FRUSZ}{14},\protect\hyperlink{ref-RJ9GQcwl}{55}{]}.
The predictions could be confirmed at least in one case, where mice received fluorescent particles during mechanical ventilation.
In 3D-visualizations the particles were predominately detected in the proximal half of the acini {[}\protect\hyperlink{ref-19jBv0ima}{56}{]}.

Based on the above-mentioned knowledge and the data of this study, we propose the following.
In the smallest acini we observed at postnatal day 4 acinar flows are mainly characterized by radial streamlines {[}\protect\hyperlink{ref-eioib1TQ}{49},\protect\hyperlink{ref-afF83siN}{57}{]}.
Diffusion governs massless particle deposition {[}\protect\hyperlink{ref-k5fqLjiy}{58}{]}.
As the acini increase in size, convective transport becomes more and more dominant in the proximal regions of the acini.
In larger acini, the proximal region shows much higher particle retention than the distal one, because in this region the larger air-flow induces a circular current, which facilitates alveolar particle deposition.
The proximal regions may even be viewed analogously to a filter capturing the particles whereas the distal regions do not receive significant amounts {[}\protect\hyperlink{ref-CVl41LwO}{12},\protect\hyperlink{ref-1DP2FRUSZ}{14},\protect\hyperlink{ref-yHHhvOtP}{15},\protect\hyperlink{ref-1HOyVjqpM}{50},\protect\hyperlink{ref-OT4s1CSX}{51},\protect\hyperlink{ref-RJ9GQcwl}{55}{]}.

@@

We thank \texttt{manubot} {[}\protect\hyperlink{ref-YuJbg3zO}{59}{]} for providing us with the means to collaboratively write and craft this manuscript.

@@

\leavevmode\hypertarget{ref-14ETRiHY2}{}%
37. Haberthür D. Data for manuscript on ``Pulmonary acini exhibit complex changes during postnatal rat lung development''. Open Science Framework. 2021. doi:\href{https://doi.org/10.17605/osf.io/jkune}{10.17605/osf.io/jkune}

\leavevmode\hypertarget{ref-pQ6Wbz73}{}%
38. Kluyver T, Ragan-Kelley B, Pérez F, Granger B, Bussonnier M, Frederic J, et al. Jupyter Notebooks -- a publishing format for reproducible computational workflows. In: Loizides F, Scmidt B, editors. IOS Press; 2016. pp. 87--90. doi:\href{https://doi.org/10.3233/978-1-61499-649-1-87}{10.3233/978-1-61499-649-1-87}

\leavevmode\hypertarget{ref-V87xbt0b}{}%
39. Haberthür D. habi/acinar-analysis: AJP-Submission. Zenodo; 2020. doi:\href{https://doi.org/10.5281/zenodo.3768407}{10.5281/zenodo.3768407}

\leavevmode\hypertarget{ref-o21zxPIu}{}%
40. Amrhein V, Greenland S, McShane B. Scientists rise up against statistical significance. Nature. 2019;567: 305--307. doi:\href{https://doi.org/10.1038/d41586-019-00857-9}{10.1038/d41586-019-00857-9}

\leavevmode\hypertarget{ref-IkHrgIj3}{}%
41. SHAPIRO SS, WILK MB. An analysis of variance test for normality (complete samples). Biometrika. 1965;52: 591--611. doi:\href{https://doi.org/10.1093/biomet/52.3-4.591}{10.1093/biomet/52.3-4.591}

\leavevmode\hypertarget{ref-MyzxBJ57}{}%
42. Mann HB, Whitney DR. On a Test of Whether one of Two Random Variables is Stochastically Larger than the Other. The Annals of Mathematical Statistics. 1947;18: 50--60. doi:\href{https://doi.org/10.1214/aoms/1177730491}{10.1214/aoms/1177730491}

\leavevmode\hypertarget{ref-ON1Bppkk}{}%
43. Kruskal WH, Wallis WA. Use of Ranks in One-Criterion Variance Analysis. Journal of the American Statistical Association. 1952;47: 583--621. doi:\href{https://doi.org/10.1080/01621459.1952.10483441}{10.1080/01621459.1952.10483441}

\leavevmode\hypertarget{ref-8Miti2Gz}{}%
44. Virtanen P, Gommers R, Oliphant TE, Haberland M, Reddy T, Cournapeau D, et al. SciPy 1.0: fundamental algorithms for scientific computing in Python. Nature Methods. 2020;17: 261--272. doi:\href{https://doi.org/10.1038/s41592-019-0686-2}{10.1038/s41592-019-0686-2}

\leavevmode\hypertarget{ref-GtCynqsI}{}%
45. Šidák Z. Rectangular Confidence Regions for the Means of Multivariate Normal Distributions. Journal of the American Statistical Association. 1967;62: 626--633. doi:\href{https://doi.org/10.1080/01621459.1967.10482935}{10.1080/01621459.1967.10482935}

\leavevmode\hypertarget{ref-I9TmP6IU}{}%
46. Zeltner TB, Bertacchini M, Messerli A, Burri PH. Morphometric Estimation of Regional Differences in the Rat Lung. Experimental Lung Research. 2009;16: 145--158. doi:\href{https://doi.org/10.3109/01902149009087879}{10.3109/01902149009087879}

\leavevmode\hypertarget{ref-Zv1XcgAO}{}%
47. Borisova E, Lovric G, Miettinen A, Fardin L, Bayat S, Larsson A, et al. Micrometer-resolution X-ray tomographic full-volume reconstruction of an intact post-mortem juvenile rat lung. Histochemistry and Cell Biology. 2020;155: 215--226. doi:\href{https://doi.org/10.1007/s00418-020-01868-8}{10.1007/s00418-020-01868-8}

\leavevmode\hypertarget{ref-AdXhaEV4}{}%
48. Mandelbrot B. How Long Is the Coast of Britain? Statistical Self-Similarity and Fractional Dimension. Science. 1967;156: 636--638. doi:\href{https://doi.org/10.1126/science.156.3775.636}{10.1126/science.156.3775.636}

\leavevmode\hypertarget{ref-eioib1TQ}{}%
49. Sznitman J, Sutter R, Altorfer D, Stampanoni M, Rösgen T, Schittny JC. Visualization of respiratory flows from 3D reconstructed alveolar airspaces using X-ray tomographic microscopy. Journal of Visualization. 2010;13: 337--345. doi:\href{https://doi.org/10.1007/s12650-010-0043-0}{10.1007/s12650-010-0043-0}

\leavevmode\hypertarget{ref-1HOyVjqpM}{}%
50. Henry FS, Haber S, Haberthür D, Filipovic N, Milasinovic D, Schittny JC, et al. The Simultaneous Role of an Alveolus as Flow Mixer and Flow Feeder for the Deposition of Inhaled Submicron Particles. Journal of Biomechanical Engineering. 2012;134: 121001. doi:\href{https://doi.org/10.1115/1.4007949}{10.1115/1.4007949}
